\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-20 chapters/covert_align.tex
Reply to DeLong test

\chapter{Compensating \acs{erp} jitter for gaze-independence}%
\label{sec:covert-align}
\epigraph{``You see, but you do not observe!''}{Sherlock Holmes in  \\ A
Scandal in Bohemia -- Sir Arthur Conan Doyle}
\noindent\emph{This chapter was published as part of~\textcite{VanDenKerchove2024}.}

@@

\textcite{Arico2014} observed higher variability in single-trial P3 peak
latencies relative to stimulus onset during covert \ac{vsa} compared to overt
\ac{vsa}\todo{qualitative, discuss}.
This latency variability contributes to reduced covert \ac{vsa} decoding performance.
While they proposed an analysis and performance prediction method, they did not provide a
decoding solution.
They suggested that compensating for latency jitter could enhance covert \ac{vsa}
decoding, but did not verify this hypothesis directly.
Additionally, \textcite{Hardiansyah2020} developed a classifier for
covert \ac{vsa} \acp{erp}, exploiting single-trial latency features in combination with
amplitude features for classification with a support vector machine.
They demonstrated the positive influence of single-trial \ac{erp} component latency
features on covert \ac{vsa} inference, yet did not attempt to correct the amplitude
features for these latencies.

@@

	\begin{subfigure}{\linewidth}
		\caption{}
		\label{fig:covert-align/block-eval}%
		\includegraphics[width=\linewidth]{figures/covert_align/figure4b.pdf}
	\end{subfigure}
  \caption[Decoder performance.]{%
		\subref{fig:covert-align/single-trial-roc-auc-diff} Difference in cross-validated
    single-trial classification \acl{rocauc} ($\Delta$ROC-AUC)
    between \ac{cble}, \ac{wcble}, and their first-stage classifier \ac{tlda}.
    95\% confidence intervals were determined using $k=1000$ bootstrapping.
    Our proposed \ac{wcble} decoder outperforms \ac{tlda} and \ac{cble} for covert and split
    ($d=3$)	\ac{vsa}. \ac{cble} scores on par with \ac{tlda}.
		\subref{fig:covert-align/block-eval} Cross-validated target selection accuracy for
    all decoders plotted as a function of the number of test repetitions in
    different VSA conditions. Significance was determined using one-sided
    (\ac{wcble} $>$ other) Wilcoxon rank-sum tests using False Discovery Rate correction
    ($*= p<0.05$, $**=p<0.01$, $***=p<0.001$). \ac{wcble} generally achieves the highest
    covert \ac{vsa} target selection accuracy.}
\end{figure}
Wilcoxon signed-rank tests controlled for multiple comparisons by Benjamini
and Hochberg's False Discovery Rate procedure (FDR) revealed that
for the BNCI2014-009 dataset, \ac{wcble} significantly outperformed \ac{tlda}
($\Delta\mathrm{ROC-AUC} = 0.019$, $p=0.004$) and \ac{cble}
($\Delta\mathrm{ROC-AUC} = 0.016$, $p=0.036$) for covert \ac{vsa},
but was significantly outperformed by \ac{tlda} in overt \ac{vsa} decoding
($\Delta\mathrm{ROC-AUC}=-0.004$, $p=0.040$).
For the CVSA-ERP dataset, \ac{wcble}
also achieved significantly better covert \ac{vsa} performance than \ac{tlda} ($\Delta\mathrm{ROC-AUC}
	= 0.023$, $p=0.041$) and \ac{cble} ($\Delta\mathrm{ROC-AUC}
	= 0.036$, $p=0.024$).
We found no significant difference in \ac{wcble} performance over \ac{tlda} in the split
\ac{vsa} conditions in the CVSA-ERP dataset, but results show a clear trend of
increased \ac{wcble} performance over \ac{tlda} and \ac{cble} as $d$ increases.
\ac{cble} failed to significantly outperform its first-stage classifier \ac{tlda} in all
evaluated \ac{vsa} conditions.
\Cref{tab:covert-align/results/single-trial} reports all single-trial classification scores for all considered models, datasets,
and conditions.

\begin{table}
  \centering
  \makebox[\textwidth][c]{%
    \input{chapters/covert_align_performance.tex}
  }
  \caption{Cross-validated single-trial classification area under the receiver operating characteristic
curve for all evaluated models, visuospatial attention conditions, and
  datasets.}
  \label{tab:covert-align/results/single-trial}
\end{table}

\Cref{fig:covert-align/block-eval} shows the cross-validated \ac{bci} selection
accuracy on the BNCI2014-009 and the CVSA-ERP datasets for all investigated
decoders.
Accuracy was determined by, for each block, selecting the character with the
highest (stage-two if applicable) classifier score and comparing it to the cued target.
Significance was calculated using one-sided Wilcoxon rank-sum tests ($p=0.05$)
corrected for FDR over decoders.
For this evaluation, no epochs were rejected to keep the trial-based structure
of \ac{bci} operation intact.
For all datasets and \ac{vsa} conditions, \ac{cble} scores approximately on par with \ac{tlda}.
Yet, \ac{wcble} yields an improved decoding accuracy for covert \ac{vsa} in both
datasets, which is greatest for smaller numbers of repetitions and
decreases as the number of repetitions increases.\todo{qualitative}
This covert \ac{vsa} accuracy increase over \ac{tlda} is significant in the
BNCI2014-009 dataset for 1 and 3 repetitions, and in CVSA-ERP for 1-5 and 10 repetitions.
Furthermore, while we reported a relative decrease in single-trial \ac{rocauc} for
\ac{wcble} in overt \ac{vsa}, this does not seem to result in a consistent decrease in target
selection accuracy.\todo{qualitative}
No significant increase of \ac{wcble} over other methods was found in split \ac{vsa}.
While Riemannian methods are significantly outperformed by \ac{tlda}, \ac{cble}, and \ac{wcble}
in the BNCI2014-009 dataset, they perform approximately on par with \ac{tlda} and
\ac{wcble} in CVSA-ERP.\todo{qualitative}

@@

  \caption[Cross-condition classifier performance and estimated jitter.]{
		\subref{fig:covert-align/covert-cross-eval},\subref{fig:covert-align/aloise2012-cross-eval}
    Difference in cross-validated area under the receiver
		operating characteristic curves between Classifier-based Latency Estimation
    (CBLE) and Woody CBLE (WCBLE) across conditions for the CVSA-ERP and
    BNCI2014-009 datasets, respectively.
    A decoder is each	time trained on a visuospatial attention (VSA) condition
    and tested on all VSA conditions.
    WCBLE yields an improvement in most
		non-overt \ac{vsa} settings, indicating it is more invariant to eye gaze than
    CBLE and \ac{tlda}.
		\subref{fig:jitter} Jitter characterized as the interquartile range (IQR)
		of target epochs for different \ac{vsa} conditions. Overt \ac{vsa} exhibits lower
    jitter than other conditions.
    Significance of differences
    was determined with two-sided Wilcoxon rank-sum tests with False Discovery
    Rate correction on per-subject jitter ($*= p<0.05$, $**=p<0.01$,
    $***=p<0.001$, $****=p<0.0001$).
	}
\end{figure*}
When comparing \ac{cble} and \ac{tlda}, we do not observe large differences in any of the
evaluated settings, similar to the within-subject conditions.\todo{qualitative}
On the contrary, when considering the comparisons between \ac{wcble} and \ac{tlda}, we
see that performance is on par or greater using \ac{wcble} for most conditions,
except for within-overt \ac{vsa} decoding.\todo{qualitative}

@@

\cref{fig:covert-align/single-trial-roc-auc-diff} shows that \ac{wcble} significantly improves
covert \ac{vsa} decoding.
This is advantageous for the development of a class of \ac{erp}-\ac{bci} interfaces for
users who prefer to rest their gaze on a fixation cross on the screen,
avoiding the effort of redirecting their eye gaze for every selection.
Furthermore, the performance gain over the first-stage classifier in split
\ac{vsa} ($d=3$) and between-\ac{vsa} condition transfer settings are promising for
users with even less eye motor control who may experience involuntary eye
movements or fixation fatigue and hence cannot keep their gaze fixed throughout an
entire \ac{bci} operation session.
\ac{wcble} would allow them to operate a \ac{bci} comfortably while directing
their gaze to whichever portion of the screen they prefer, even when there is
another target present or this location varies during the
course of operation.
Although \ac{wcble} did not significantly improve overt \ac{vsa} single-trial decoding,
\cref{fig:covert-align/block-eval} shows that this does not negatively impact target
selection accuracy.
While target selection accuracy also did not improve for split \ac{vsa}, the
increase in single-trial performance in split ($d=3$) shows that an iterative
alignment procedure has the potential to improve over \ac{cble} and its first-stage
classifier in this case as well.

We believe the relative increase in performance of our proposed decoder in covert and
split \ac{vsa}, and the lack thereof in overt \ac{vsa}, could stem from the following:
\begin{enumerate*}[label=(\arabic*)]
  \item Covert and split \ac{vsa} exhibit higher P3 jitter than overt \ac{vsa}.
        In covert and split \ac{vsa}, participants have to execute a dual task by
        dissociating their visuospatial attention and gaze fixation.
        Evidence shows that \ac{erp} latency variability is
        higher when attention is
        divided~\cite{Polich2007,Arico2014}.
        \textcite{Arico2014} also partly attribute higher latency jitter to the covert \ac{vsa}
        task performed in the BNCI2014-009 dataset, since the GeoSpell interface
        requires both spatial and feature attention.
  \item In overt \ac{vsa}, the first-stage classifier can rely mostly on the modulation
        of early visually evoked potentials (VEPs) like N1 rather than on the
        P3~\cite{Treder2010}.
        These VEPs are closely related to visual processing, hence exhibit
        lower jitter, contrary to P3, which is more prone to the effects of
        attention and workload~\cite{Hu2010}, reducing the contribution of
        alignment.
  \item This property can also result in the estimation of VEP latencies
        instead of the P3 latency, and \ac{wcble} would in this case fail to
        increase the P3 \ac{snr}, which still could be somewhat jittered in overt
        \ac{vsa}.
  \item Aligning to the P3 will lower the \ac{snr} of early VEPs, while aligning to
        VEPs will lower P3 \ac{snr}, since they are not time-locked to each other.
	\item Covert and split \ac{vsa} \acp{erp} may exhibit lower \ac{snr} than overt
        \ac{vsa} due to lower P3 amplitudes or even due to the presence of higher P3
        jitter itself.
        Higher \ac{snr} in overt \ac{vsa} results in higher decoding performance of
        state-of-the-art classifiers, leaving less room for relative improvement
        in this case.
\end{enumerate*}
\todo{first item qualitative}

Although it is not immediately clear if \ac{wcble} actively corrects for higher P3 jitter
present in covert and split \ac{vsa} compared to overt \ac{vsa},
we justify our approach in a similar manner to \textcite{Hardiansyah2020} by
observing that the increased discrimination performance of a machine-learning
model accounting for jitter forgoes the need for characterizing the underlying
physiological processes, while still objectively quantifying the presence of
jitter between data classes.
Furthermore, some evidence points toward higher P3 jitter as the main
contributing factor.
While \autoref{fig:erps} shows no visible smearing effect in
the shape of the \acp{erp}, a more quantitative analysis on latencies in
\autoref{fig:jitter} indicates the opposite.
To obtain comparable results across \ac{vsa} conditions, \ac{wcble} was
evaluated per session and trained on all combined \ac{vsa} conditions as in
\cref{sec:covert-align/results/cross}.
Since all conditions have the P3 in common, the estimated latencies can be
interpreted as P3 latencies.
Two-sided Wilcoxon rank-sum tests on jitter, expressed as the interquartile
range (IQR) of the estimated latencies of target trials, revealed that overt
\ac{vsa} exhibited significantly lower jitter than all other conditions for
both datasets.\todo{qualitative}
No other significant differences were found.
$p$-values were corrected for the family-wise error rate using Bonferroni
correction.
Additionally, \textcite{Arico2014} prove that P3 jitter does play a
non-negligible role in covert \ac{vsa} by comparing performance between
overt and covert \ac{vsa}, while including or excluding early VEPs.
This analysis showed that the absence of the N1 and other early VEPs is not the
only factor hampering covert \ac{vsa} decoding performance.
However, the large increase in performance for covert \ac{vsa} reported by
\textcite{Arico2014}\todo{qualitative} could also
be explained by overfitting on artifacts amplified by aligning, since their
method is not evaluated on unseen data, as opposed to ours.\todo{qualitative}

@@

>>>>

\chapter{Compensating \acs{erp} jitter for gaze-independence}%
\label{sec:covert-align}
\epigraph{``You see, but you do not observe!''}{Sherlock Holmes in  \\ \emph{A
Scandal in Bohemia -- Sir Arthur Conan Doyle}}
\noindent\emph{This chapter was published as part of~\textcite{VanDenKerchove2024}.}

@@

\textcite{Arico2014} observed higher variability in single-trial P3 peak
latencies relative to stimulus onset during covert \ac{vsa} (529 ms$^2$) compared to overt
\ac{vsa} (256$^2$ ms).
This latency variability contributes to reduced covert \ac{vsa} decoding performance.
While they proposed an analysis and performance prediction method, they did not provide a
decoding solution.
They suggested that compensating for latency jitter could enhance covert \ac{vsa}
decoding, but did not verify this hypothesis directly.
Additionally, \textcite{Hardiansyah2020} developed a classifier for
covert \ac{vsa} \acp{erp}, exploiting single-trial latency features in combination with
amplitude features for classification with a support vector machine.
They demonstrated the positive influence of single-trial \ac{erp} component latency
features on covert \ac{vsa} inference, yet did not attempt to correct the amplitude
features for these latencies.

@@

	\begin{subfigure}{\linewidth}
		\caption{}
		\label{fig:covert-align/block-eval}%
		\includegraphics[width=\linewidth]{figures/covert_align/figure4b.pdf}
	\end{subfigure}
  \caption[Decoder performance.]{%
		\subref{fig:covert-align/single-trial-roc-auc-diff} Difference in cross-validated
    single-trial classification \acl{rocauc} ($\Delta$ROC-AUC)
    between \ac{cble}, \ac{wcble}, and their first-stage classifier \ac{tlda}.
    95\% confidence intervals were determined using $k=1000$ bootstrapping.
    Our proposed \ac{wcble} decoder outperforms \ac{tlda} and \ac{cble} for covert and split
    ($d=3$)	\ac{vsa}. \ac{cble} scores on par with \ac{tlda}.
		\subref{fig:covert-align/block-eval} Cross-validated target selection accuracy for
    all decoders plotted as a function of the number of test repetitions in
    different VSA conditions. Significance was determined using one-sided
    (\ac{wcble} $>$ other) Wilcoxon signed-rank tests using False Discovery Rate correction
    ($*= p<0.05$, $**=p<0.01$, $***=p<0.001$). \ac{wcble} generally achieves the highest
    covert \ac{vsa} target selection accuracy.}
\end{figure}

Wilcoxon signed-rank tests controlled for multiple comparisons by Benjamini
and Hochberg's False Discovery Rate procedure (FDR) revealed that
for the BNCI2014-009 dataset, \ac{wcble} significantly outperformed \ac{tlda}
($\Delta\mathrm{ROC-AUC} = 0.019$, $p=0.004$) and \ac{cble}
($\Delta\mathrm{ROC-AUC} = 0.016$, $p=0.036$) for covert \ac{vsa},
but was significantly outperformed by \ac{tlda} in overt \ac{vsa} decoding
($\Delta\mathrm{ROC-AUC}=-0.004$, $p=0.040$).
For the CVSA-ERP dataset, \ac{wcble}
also achieved significantly better covert \ac{vsa} performance than \ac{tlda} ($\Delta\mathrm{ROC-AUC}
	= 0.023$, $p=0.041$) and \ac{cble} ($\Delta\mathrm{ROC-AUC}
	= 0.036$, $p=0.024$).
We found no significant difference in \ac{wcble} performance over \ac{tlda} in the split
\ac{vsa} conditions in the CVSA-ERP dataset, but results show a clear trend of
increased \ac{wcble} performance over \ac{tlda} and \ac{cble} as $d$ increases.
\ac{cble} failed to significantly outperform its first-stage classifier \ac{tlda} in all
evaluated \ac{vsa} conditions.
\Cref{tab:covert-align/results/single-trial} reports all single-trial classification scores for all considered models, datasets, and conditions.

\Cref{fig:covert-align/block-eval} shows the cross-validated \ac{bci} selection
accuracy on the BNCI2014-009 and the CVSA-ERP datasets for all investigated
decoders.
Accuracy was determined by, for each block, selecting the character with the
highest (stage-two if applicable) classifier score and comparing it to the cued target.
Significance was calculated using one-sided Wilcoxon signed-rank tests ($p=0.05$)
corrected for FDR over decoders.
For this evaluation, no epochs were rejected to keep the trial-based structure
of \ac{bci} operation intact.
For all datasets and \ac{vsa} conditions, \ac{cble} scores approximately on par with \ac{tlda}.
Yet, \ac{wcble} yields an improved decoding accuracy for covert \ac{vsa} in both
datasets, which is greatest for smaller numbers of repetitions and
decreases as the number of repetitions increases.
This covert \ac{vsa} accuracy increase over \ac{tlda} is significant in the
BNCI2014-009 dataset for 1 and 3 repetitions, and in CVSA-ERP for 1-5 and 10 repetitions.
Furthermore, while we reported a relative decrease in single-trial \ac{rocauc} for
\ac{wcble} in overt \ac{vsa}, this does not seem to result in a consistent decrease in target
selection accuracy.
No significant increase of \ac{wcble} over other methods was found in split \ac{vsa}.
While Riemannian methods are significantly outperformed by \ac{tlda}, \ac{cble}, and \ac{wcble}
in the BNCI2014-009 dataset, they perform approximately on par with \ac{tlda} and
\ac{wcble} in CVSA-ERP.
\todo{qualitative, list accuracies for one repetition}

@@

  \caption[Cross-condition classifier performance and estimated jitter.]{
		\subref{fig:covert-align/covert-cross-eval},\subref{fig:covert-align/aloise2012-cross-eval}
    Difference in cross-validated area under the receiver
		operating characteristic curves between Classifier-based Latency Estimation
    (CBLE) and Woody CBLE (WCBLE) across conditions for the CVSA-ERP and
    BNCI2014-009 datasets, respectively.
    A decoder is each	time trained on a visuospatial attention (VSA) condition
    and tested on all VSA conditions.
    WCBLE yields an improvement in most
		non-overt \ac{vsa} settings, indicating it is more invariant to eye gaze than
    CBLE and \ac{tlda}.
		\subref{fig:jitter} Jitter characterized as the interquartile range (IQR)
		of target epochs for different \ac{vsa} conditions. Overt \ac{vsa} exhibits lower
    jitter than other conditions.
    Significance of differences
    was determined with two-sided Wilcoxon signed-rank tests with False Discovery
    Rate correction on per-subject jitter ($*= p<0.05$, $**=p<0.01$,
    $***=p<0.001$, $****=p<0.0001$).
	}
\end{figure*}
When comparing \ac{cble} and \ac{tlda}, we do not observe large differences in any of the
evaluated settings, similar to the within-subject conditions.
\todo{qualitative}
On the contrary, when considering the comparisons between \ac{wcble} and \ac{tlda}, we
see that performance is on par or greater using \ac{wcble} for most conditions,
except for within-overt \ac{vsa} decoding.\todo{qualitative}

@@

\Cref{fig:covert-align/single-trial-roc-auc-diff} shows that \ac{wcble} significantly improves
covert \ac{vsa} decoding.
This is advantageous for the development of a class of \ac{erp}-\ac{bci} interfaces for
users who prefer to rest their gaze on a fixation cross on the screen,
avoiding the effort of redirecting their eye gaze for every selection.
Furthermore, the performance gain over the first-stage classifier in split
\ac{vsa} ($d=3$) and between-\ac{vsa} condition transfer settings are promising for
users with even less eye motor control who may experience involuntary eye
movements or fixation fatigue and hence cannot keep their gaze fixed throughout an
entire \ac{bci} operation session.
\ac{wcble} would allow them to operate a \ac{bci} comfortably while directing
their gaze to whichever portion of the screen they prefer, even when there is
another target present or this location varies during the
course of operation.
Although \ac{wcble} did not significantly improve overt \ac{vsa} single-trial decoding,
\cref{fig:covert-align/block-eval} shows that this does not negatively impact target
selection accuracy.
While target selection accuracy also did not improve for split \ac{vsa}, the
increase in single-trial performance in split ($d=3$) shows that an iterative
alignment procedure has the potential to improve over \ac{cble} and its first-stage
classifier in this case as well.

We believe the relative increase in performance of our proposed decoder in covert and
split \ac{vsa}, and the lack thereof in overt \ac{vsa}, could stem from the following:
\begin{enumerate*}[label=(\arabic*)]
  \item Covert and split \ac{vsa} exhibit higher P3 jitter than overt \ac{vsa}.
        In covert and split \ac{vsa}, participants have to execute a dual task by
        dissociating their visuospatial attention and gaze fixation.
        Evidence shows that \ac{erp} latency variability is
        higher when attention is divided~\cite{Polich2007,Arico2014}.
        \textcite{Arico2014} also partly attribute higher latency jitter to the covert \ac{vsa}
        task performed in the BNCI2014-009 dataset, since the GeoSpell interface
        requires both spatial and feature attention.
  \item In overt \ac{vsa}, the first-stage classifier can rely mostly on the modulation
        of early visually evoked potentials (VEPs) like N1 rather than on the
        P3~\cite{Treder2010}.
        These VEPs are closely related to visual processing, hence exhibit
        lower jitter, contrary to P3, which is more prone to the effects of
        attention and workload~\cite{Hu2010}, reducing the contribution of
        alignment.
  \item This property can also result in the estimation of VEP latencies
        instead of the P3 latency, and \ac{wcble} would in this case fail to
        increase the P3 \ac{snr}, which still could be somewhat jittered in overt
        \ac{vsa}.
  \item Aligning to the P3 will lower the \ac{snr} of early VEPs, while aligning to
        VEPs will lower P3 \ac{snr}, since they are not time-locked to each other.
	\item Covert and split \ac{vsa} \acp{erp} may exhibit lower \ac{snr} than overt
        \ac{vsa} due to lower P3 amplitudes or even due to the presence of higher P3
        jitter itself.
        Higher \ac{snr} in overt \ac{vsa} results in higher decoding performance of
        state-of-the-art classifiers, leaving less room for relative improvement
        in this case.
\end{enumerate*}

Although it is not immediately clear if \ac{wcble} actively corrects for higher P3 jitter
present in covert and split \ac{vsa} compared to overt \ac{vsa},
we justify our approach in a similar manner to \textcite{Hardiansyah2020} by
observing that the increased discrimination performance of a machine-learning
model accounting for jitter forgoes the need for characterizing the underlying
physiological processes, while still objectively quantifying the presence of
jitter between data classes.
Furthermore, some evidence points toward higher P3 jitter as the main
contributing factor.
While \autoref{fig:erps} shows no visible smearing effect in
the shape of the \acp{erp}, a more quantitative analysis on latencies in
\autoref{fig:jitter} indicates the opposite.
To obtain comparable results across \ac{vsa} conditions, \ac{wcble} was
evaluated per session and trained on all combined \ac{vsa} conditions as in
\cref{sec:covert-align/results/cross}.
Since all conditions have the P3 in common, the estimated latencies can be
interpreted as P3 latencies.
Two-sided Wilcoxon rank-sum tests on jitter, expressed as the interquartile
range (IQR) of the estimated latencies of target trials, revealed that overt
\ac{vsa} exhibited significantly lower jitter than all other conditions for
both datasets.\todo{qualitative}
No other significant differences were found.
$p$-values were corrected for the family-wise error rate using Bonferroni
correction.
Additionally, \textcite{Arico2014} prove that P3 jitter does play a
non-negligible role in covert \ac{vsa} by comparing performance between
overt and covert \ac{vsa}, while including or excluding early VEPs.
This analysis showed that the absence of the N1 and other early VEPs is not the
only factor hampering covert \ac{vsa} decoding performance.
However, the large increase in accuracy for covert \ac{vsa} reported by
\textcite{Arico2014} (60\% to 95\% with a window focusing on only the P3
component) could also be explained by overfitting on artifacts amplified by
aligning, since their method is not evaluated on unseen data, as opposed to ours.

@@

\clearpage
\begin{subappendices}
  \section{Single-trial classification performances}
  \begin{table}[ht]
  \centering
  \makebox[\textwidth][c]{%
    \input{chapters/covert_align_performance.tex}
  }
  \caption{Cross-validated single-trial classification area under the receiver operating characteristic
curve for all evaluated models, visuospatial attention conditions, and
  datasets.}
  \label{tab:covert-align/results/single-trial}
\end{table}

  \clearpage
  \section{Cross-condition transfer classification performances}

  \begin{table}[ht]
      \footnotesize
      \input{chapters/covert_align_cross-CVSA-ERP.tex}
      \caption[Transfer learning results for CVSA-ERP]{Cross-condition transfer
      classification performance (cross-validated \ac{rocauc})
      for all evaluated models in the CVSA-ERP dataset. Each decoder is trained
      on one \ac{vsa} condition and tested across all
      \ac{vsa}
      conditions to evaluate gaze-independence and transfer learning
      performance.}
  \end{table}
  \begin{table}[ht]
      \footnotesize
      \input{chapters/covert_align_cross-BNCI2014-009.tex}
      \caption[Transfer learning results for BNCI2014-009]{Cross-condition transfer
      classification performance (cross-validated \ac{rocauc})
      for all evaluated models in the CVSA-ERP dataset. Each decoder is trained
      on one \ac{vsa} condition and tested across all \ac{vsa}
      conditions to evaluate gaze-independence and transfer learning
      performance.}
  \end{table}
\end{subappendices}

### 2024-11-20 chapters/patients.tex
Reply to DeLong test

\chapter{Case studies}
\label{sec:patients}
\section{Introduction}

@@

However, \ac{wcble} did not improve \ac{tlda} performance in the free \ac{vsa} setting, but
XDAWNCov+TS+LDA performance was slightly lower here (though not
significantly).
(\ac{wcble} 74.15\%, XDAWNCov+TS+LDA 71.88\%, \ac{tlda} 74.27\%).
More interestingly, we noticed that performances of the decoders in free
\ac{vsa} were close to those in the overt \ac{vsa}.
A substantial decrease in performance from the overt setting to the free
setting was observed for
subjects PC3 and PC4.
For PB2, who also relied on covert \ac{vsa} during the uncued free \ac{vsa}
setting, there was no decrease in performance.

>>>>

\chapter{Case studies}
\label{sec:patients}
\epigraph{I knew there would be a lot of unexpected issues involved with
patient experiments, but I never thought there would be a baby
goats in the experiment room.}{Juliette}

\section{Introduction}

@@

However, \ac{wcble} did not improve \ac{tlda} performance in the free \ac{vsa} setting, but
XDAWNCov+TS+LDA performance was slightly lower here (though not
significantly).
(\ac{wcble} 74.15\%, XDAWNCov+TS+LDA 71.88\%, \ac{tlda} 74.27\%).
More interestingly, we noticed that performances of the decoders in free
\ac{vsa} were close to those in the overt \ac{vsa}.
A substantial decrease in performance from the overt setting to the free
setting was observed for subjects PC3
(\ac{wcble}: $70.31\%>62.14\%$, XDAWNCov+TS+LDA: $65.78\%>62.18\%$, \ac{tlda}:
$70.49\%>63.76\%$)
and PC4
(\ac{wcble}: $65.56\%>55.71\%$, XDAWNCov+TS+LDA: $62.02\%>54.24\%$, \ac{tlda}:
$66.12\%>57.08\%$).
For PB2, who also relied on covert \ac{vsa} during the uncued free \ac{vsa}
according to gaze tracking setting, the decrease in performance was also
present, but not as substantial
(\ac{wcble}: $82.76\%>78.88\%$, XDAWNCov+TS+LDA: $80.74\%>77.99\%$, \ac{tlda}:
$83.21\%>78.84\%$).

### 2024-11-20 chapters/stbf_struct.tex
Quantify qualitative statements in bttda chapter
Reply to DeLong test

	\subsection{Classifier accuracy for limited training data}
	It is of interest to keep the calibration time before \ac{bci}
	operation as short as possible.
	We mimic this problem by training the classifier with as few training epochs as possible.
	We evaluate the performance of all classifiers for different levels of
	available training data and apply the cross-validation procedure nine times (the number of blocks in the training fold) for all subjects, keeping the
	corresponding number of blocks in the training folds and dropping the rest.
  \Cref{fig:accuracy}, \cref{fig:stbf-struct/accuracy-ap} and
  \crefrange{tab:stbf-struct/accuracy/stbf-struct}{tab:stbf-struct/accuracy/xdawn-rg}
  shows each classifier's accuracy relative to the data availability and trials
  used for agveraging.
	We statistically compare the two newly proposed classifiers,
	\textsc{stbf-struct} and \textsc{stbf-shrunk} with different levels of training
	data availability using a one-sided paired Wilcoxon rank-sum test with Holm correction for the multiple pairwise comparisons between classifiers.
	Statistical comparisons were performed for trial
  \cref{tab:stbf-struct/p-values-1}, two \cref{tab:stbf-struct/p-values-2}
  and five trials \cref{tab:stbf-struct/p-values-5}.

@@

  \footnotesize
  \begin{table}[!h]
	    \input{chapters/stbf_struct-acc-stbf_kron_toep_shrunk.tex}
      \caption[Accuracies for \textsc{stbf-struct}.]{Accuracies (\%) for \textsc{stbf-struct}.}
      \label{tab:stbf-struct/accuracies/stbf-struct}
  \end{table}
  \begin{table}[!h]
	    \input{chapters/stbf_struct-acc-stbf_loocv.tex}
      \caption[Accuracies for \textsc{stbf-struct}.]{Accuracies (\%) for
      \textsc{stbf-shrunk}.}
      \label{tab:stbf-struct/accuracies/stbf-shrunk}
  \end{table}
  \begin{table}[!h]
	    \input{chapters/stbf_struct-acc-stbf_emp.tex}
      \caption[Accuracies for \textsc{stbf-struct}.]{Accuracies (\%) for
      \textsc{stbf-emp}.}
      \label{tab:stbf-struct/accuracies/stbf-emp}
  \end{table}
  \begin{table}[!h]
	    \input{chapters/stbf_struct-acc-xdawn_rg.tex}
      \caption[Accuracies for \textsc{stbf-struct}.]{Accuracies (\%) for
      \textsc{xdawn+rg}.}
      \label{tab:stbf-struct/accuracies/xdawn-rg}
  \end{table}

@@

    \begin{subtable}{\linewidth}
	    \input{chapters/stbf_struct_tests_3.tex}
	    \caption{Evaluation averaging over five testing trials.}
	    \label{tab:stbf-struct/p-values-5}
    \end{subtable}
    \caption[Statistical significance of differences in classifier
    performance.]{%
        Statistical significance of differences in classifier performance.
        $p$-values calculated by one-sided paired Wilcoxon rank-sum test with
	    	Holm correction using one testing trial for different classifiers
	    	and levels of data availability. $p$-values $<0.05$ are considered
        significant.
    }
	  \label{tab:stbf-struct/p-values}
	\end{table}

>>>>

	\subsection{Classifier accuracy for limited training data}
	It is of interest to keep the calibration time before \ac{bci}
	operation as short as possible.
	We mimic this problem by training the classifier with as few training epochs as possible.
	We evaluate the performance of all classifiers for different levels of
	available training data and apply the cross-validation procedure nine times (the number of blocks in the training fold) for all subjects, keeping the
	corresponding number of blocks in the training folds and dropping the rest.
  \Cref{fig:accuracy}, \cref{fig:stbf-struct/accuracy-ap} and
  \crefrange{tab:stbf-struct/accuracies/stbf-struct}{tab:stbf-struct/accuracies/xdawn-rg}
  shows each classifier's accuracy relative to the data availability and trials
  used for agveraging.
	We statistically compare the two newly proposed classifiers,
	\textsc{stbf-struct} and \textsc{stbf-shrunk} with different levels of training
	data availability using a one-sided Wilcoxon signed-rank test with Holm correction for the multiple pairwise comparisons between classifiers.
	Statistical comparisons were performed for trial
  \cref{tab:stbf-struct/p-values-1}, two \cref{tab:stbf-struct/p-values-2}
  and five trials \cref{tab:stbf-struct/p-values-5}.

@@

  \begin{table}[!h]
      \footnotesize
	    \input{chapters/stbf_struct-acc-stbf_kron_toep_shrunk.tex}
      \caption[Accuracies for \textsc{stbf-struct}.]{Accuracies (\%) for \textsc{stbf-struct}.}
      \label{tab:stbf-struct/accuracies/stbf-struct}
  \end{table}
  \begin{table}[!h]
      \footnotesize
	    \input{chapters/stbf_struct-acc-stbf_loocv.tex}
      \caption[Accuracies for \textsc{stbf-struct}.]{Accuracies (\%) for
      \textsc{stbf-shrunk}.}
      \label{tab:stbf-struct/accuracies/stbf-shrunk}
  \end{table}
  \begin{table}[!h]
      \footnotesize
	    \input{chapters/stbf_struct-acc-stbf_emp.tex}
      \caption[Accuracies for \textsc{stbf-struct}.]{Accuracies (\%) for
      \textsc{stbf-emp}.}
      \label{tab:stbf-struct/accuracies/stbf-emp}
  \end{table}
  \begin{table}[!h]
      \footnotesize
	    \input{chapters/stbf_struct-acc-xdawn_rg.tex}
      \caption[Accuracies for \textsc{stbf-struct}.]{Accuracies (\%) for
      \textsc{xdawn+rg}.}
      \label{tab:stbf-struct/accuracies/xdawn-rg}
  \end{table}

@@

    \begin{subtable}{\linewidth}
	    \input{chapters/stbf_struct_tests_3.tex}
	    \caption{Evaluation averaging over five testing trials.}
	    \label{tab:stbf-struct/p-values-5}
    \end{subtable}
    \caption[Statistical significance of differences in classifier
    performance.]{%
        Statistical significance of differences in classifier performance.
        $p$-values calculated by one-sided Wilcoxon signed-rank test with
	    	Holm correction using one testing trial for different classifiers
	    	and levels of data availability. $p$-values $<0.05$ are considered
        significant.
    }
	  \label{tab:stbf-struct/p-values}
	\end{table}
